Staged Abstract Interpreters

### 2018-10-27 paper/10_intro.tex
revise on intro

2. Multi-stage programming is widely used in high performance computation.
Precise whole-program analysis can be very slow, how to optimize it systematically/mechanically?
bring existing high-performance computation facilities, in the form of meta-programming, into
semantic artifacts of abstract interpretation, i.e., abstract interpreters.
and bridging the gap between designing sound analyses and implementing efficient analyses.
without changing the abstract interpreter, just type

partial evaluation on program analyzers is not a new idea, remove the interpretive overhead of repetively 
traversing the abstract syntax tree.
we revisit this approach from a more recent perspective by combining general abstract interpreters
and multi-stage programming.

When staging a concrete interpreter, the programmers need to distinguish static and dynamic values
--- the given program is classified as static because it is known at compile-time, and the inputs to that program 
are dynamic. However, when staging an abstract interpreter, this distinction not exists anymore. 
Because an abstract interpreter treats all the inputs as some form of abstract values, which usually are 
top elements in their abstract domains and are statically known. Then what is the point of staging?
A surprising by-product of thinking about this question is to realize that we can apply the staged
abstract interpreter on \textit{open} programs, and the free variables representing other parts of 
the program (e.g., libraries) are dynamic input, then we obtain a modular analysis, mechanically.

@@

\begin{itemize}
\item Starting from a definitional abstract interpreter for a higher-order functional language,
  we show how to turn it into a staged abstract interpreter by adding stage annotations with 
  Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
  We later extend the abstract interpreter to support
  imperative language features, such as states, loops and exceptions.
\item We elaborate the idea that the staged abstract interpreter not only improves the performance 
  but also the scalability, which are two major issues in static analysis, if we apply it on open
  programs. TODO: more on this point.
\item To evaluate the generality of the idea, we conduct several case studies. 
  First we revisited Abstract Compilation (TODO: cite), which is a closure generation technique applied 
  on 0-CFA and can be understood as an instance of staged abstract interpreter. 
  Second, we demonstrate existing control-flow analysis techniques, such as context/flow/path-sensitive 
  analysis and store-widening can be integrated with staged abstract interpreters.
  Finally, we show it also can be applied to numerical analysis in an imperative language.
  (Optional: taint/information flow analysis?)
\item We also emprically evaluate the performance improved of staged abstract interpreter with the
  unstaged version and other existing optimizations on benchmarks.
\end{itemize}

>>>>

(TODO: more intro on MSP)
Multi-stage programming has been widely used to improve the performance in many applications, 
such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
generic programming \cite{DBLP:journals/pacmpl/Yallop17}, 
data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
query compilation of databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
and etc.
Specializing static analyses by partial evaluation had been emerged in 90s \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587}, 
and indeed it is able to effectively remove the interpretive overhead of repetively traversing the abstract syntax tree.
However, it is still worth to revisit the idea from a modern perspective based on meta-programming for general abstract interpreters.
Practically, precise whole-program analyses can be very slow due to the inherent complexity of the algorithm and the large scale of the analyzed programs (TODO: cite); 
particularly, we would like a systematic approach to optimize program analyzers and meanwhile minimally modifying the analyzer programs.
In this paper, we show how to achieve this goal by bring existing high-performance computing facilities, in the form of generative
programming, to abstract interpretation semantic artifacts, i.e., abstract interpreters. We use LMS framework, which enables
multi-stage programming by just changing the types of ther analyzer program, and we thus bridge the gap between designing sound 
program analyses and implementing efficient analyzers.

When staging a concrete interpreter, the programmers need to distinguish static and dynamic values
--- the given program to be executed by the interpreter is classified as static because it is known 
at compile-time, and the inputs to that program are dynamic. 
However, when staging an abstract interpreter, this distinction not exists anymore.  
Because the abstract interpreter instantiates all the inputs as some form of abstract values, which are usually
top elements in their abstract domains and are statically known. 
Then what is the point of staging if no such distinction?
A surprising by-product of thinking about this question is to realize that we can apply the staged
abstract interpreter on \textit{open} programs, and the free variables representing other parts of 
the program (e.g., libraries) are dynamic input, then we obtain a modular analysis, mechanically.
(TODO: unknown context)

@@

\begin{itemize}
\item Starting from a definitional abstract interpreter for a higher-order functional language,
  which the syntax and semantics are described in section \ref{bg_lang},
  we show how to turn it into a staged abstract interpreter by adding stage annotations within
  Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
  We later extend the staged abstract interpreter to support imperative language features 
  (section \ref{cases_imp}), such as states, loops and exceptions.
\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{performance}
  but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
  programs. TODO: more on this point.
\item To evaluate the generality and applicability of the idea, we conduct three case studies in section \ref{cases_study}. 
  First we revisited Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
  technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 
  staged abstract interpreter. 
  Second, we demonstrate sophisticated control-flow analysis techniques, such as adaptive context/flow/path-sensitive 
  analysis and store-widening can be integrated with staged abstract interpreters without losing any precision.
  Finally, we show that in an imperative language, the idea can be applied onto an abstract interpreter with
  an abstract domain that performs numerical analysis.
  (Optional: taint/information flow analysis?)
\item We also emprically evaluate the performance improved by staging an big-step abstract interpreter 
  for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 
  and an optimizing small-step version.
\end{itemize}

### 2018-10-27 paper/abstract.tex
revise on intro

In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
staged abstract interpreters. 
By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
By appying the abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
we obtain a modular analysis that generates sound partial analysis results which can be composed 
and reused later without losing precision, though the original abstract interpreter is a whole-program 
analysis algorithm.

>>>>

In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
staged abstract interpreters. 
By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
By appying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
we obtain a modular analysis that generates sound partial analysis results which can be composed 
and reused later without losing precision, though the original abstract interpreter is a whole-program 
analysis algorithm.
